---
title: "Exploratory Data Analysis"
author: "Kevin Thompson, Sean Kennedy, Sachin Chavan"
date: "September 29, 2019"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(ggcorrplot)
library(visdat)
library(corrplot)
library(ggthemes)
library(sqldf)
library(rjson)
library(purrr)
```

## 1.1: Load data
```{r color="black"}
load("../analysis/data/.RData")
#load("../analysis/data/RData.dms") 
QBCrossSectional$Week <- as.factor(QBCrossSectional$Week)
attach(QBCrossSectional)
```

## 1.2: Structures QBCrossSectional and QBPanels

### 1.2.1 QB (Quarterback)
```{r}
head(QBCrossSectional)
summary(QBCrossSectional)
```
### 1.2.2 QB (Panel Data)
```{r}
str(QBPanels)
```

## 1.3: Missing Data 

### 1.3.1 Quarterbacks (Old Data QBdata dataset)

```{r}
vis_dat(QBCrossSectional[1:30,])
vis_dat(QBCrossSectional[31:60, ])
vis_dat(QBCrossSectional[61:89, ])
```

### 1.3.2 QBCrossSectional dataset (New dataset)
```{r}
vis_dat(QBCrossSectional)
print(QBCrossSectional)
```

###  1.3.3 QBPanels dataset (New dataset)
```{r}
print(QBPanels)

for (panel in 1:length(QBPanels)) {
   missing_plot <- vis_miss(as.data.frame(QBPanels[panel]))
   print(missing_plot)
}
```

## 1.4: Corrrelogram 
### 1.4.1 Corrrelogram Style1 QBData (Old dataset)

```{r}
#QBdata_num <- QBCrossSectional[-c(1:3,5:16,29:36,39:52,55:56,74:89)]
#corr <- round(cor(QBdata_num), 1)

#ggcorrplot(corr, hc.order = TRUE, 
#           type = "full", 
#           lab = TRUE, 
#           lab_size = 1.5, 
#           method="square", 
#           colors = c("tomato2", "white", "springgreen3"), 
#           title="Correlogram of Quarterbacks", tl.cex = 7,pch=2,pch.col =3,show.diag = T,
#           ggtheme=theme_classic)

```

### 1.4.2 Corrrelogram Style2 QBData (Old dataset)

```{r}
#corrplot(corr, method="circle",tl.cex = 0.55,tl.col = "#1C1C1C")
```

### 1.4.3 Corrrelogram of Crosssectional data (New dataset)

```{r}
QBX_ds <- QBCrossSectional[-c(1:6,21)]
corr <- round(cor(QBX_ds), 1)
ggcorrplot(corr, hc.order = TRUE, 
           type = "full", 
           lab = TRUE, 
           lab_size = 1.5, 
           method="square", 
           colors = c("tomato2", "white", "springgreen3"), 
           title="Correlogram of Quarterbacks", tl.cex = 7,pch=2,pch.col =3,show.diag = T,
           ggtheme=theme_classic)
```

## 1.5: Distributions
### 1.5.1 Density plot for Fantasypoints is approxmiately Normal (Old QBData)

```{r}
library(e1071)
par(mfrow=c(1, 2))  # divide graph area in 2 columns
target <- QBCrossSectional$FantasyPoints
plot(density(target), main="QBdata: FantasyPoints", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(target), 2)))
polygon(density(target), col="red",asp=1.2 )

```

### 1.5.2 Density plot for Fantasypoints is approxmiately Normal (New QBCrossSectional)

```{r}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
target <- QBCrossSectional$FantasyPoints
plot(density(target), main="CrossSectional Dataset: FantasyPoints", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(target), 2)))
polygon(density(target), col="red")

```

### 1.5.3 Boxplots - Target and Individual Predictor Bevavior for per Team

I think we should do these by week - not team - we want to look at variance of the time series across obervations
```{r}

QBCrossSectional %>% ggplot(aes(y=FantasyPoints,x=Week,fill=Week,group=Week))+
                     geom_boxplot(show.legend = FALSE)+
                     xlab("Week")+ylab("FantasyPoints")+
                     labs(title="Fantasyfootball", 
                     subtitle="Weekly fantasypoints", 
                     aption="Source: Fantasyfootball")+theme_economist()

for (i in 7:19) {

ggplotp <- QBCrossSectional %>% ggplot(aes_string(y=names(QBCrossSectional[i]),x=Week,fill=Week,group=Week))+
                     geom_boxplot(show.legend = FALSE)+
                     xlab("Week")+ylab(names(QBCrossSectional[i]))+
                     labs(title="Fantasyfootball", 
                     subtitle="Weekly fantasypoints", 
                     aption="Source: Fantasyfootball")+theme_economist()  
  print(ggplotp)
}



```

## 1.6: Relationships
### 1.6.1 Relationship between FanatasyPoints with all predictors and taking Opponent into consideration

Not sure that combining these is useful

```{r}
attach(QBCrossSectional)
QBCrossSectional %>% ggplot(aes(y=FantasyPoints,x=PassingCompletions+PassingAttempts+
                                                  PassingCompletionPercentage+
                                                  PassingYards+PassingYardsPerAttempt+
                                                  PassingTouchdowns+PassingInterceptions+
                                                  PassingRating+RushingAttempts+RushingYards+
                                                  RushingYardsPerAttempt+RushingTouchdowns+FumblesLost,
                                                  color=Opponent))+ 
                     xlab("Relationship of predictors with each Opponent")+
                     geom_point()+
                     geom_smooth(method="lm",se=F)+
                     theme(axis.text.x = element_text(angle=65, vjust=0.6))+facet_wrap(~Opponent)

``` 

### 1.6.2 Relationship between FanatasyPoints with all predictors and taking Home turf into consideration

Same here.....
```{r}
attach(QBCrossSectional)
QBCrossSectional %>% ggplot(aes(y=FantasyPoints,x=PassingCompletions+PassingAttempts+
                                                  PassingCompletionPercentage+
                                                  PassingYards+PassingYardsPerAttempt+
                                                  PassingTouchdowns+PassingInterceptions+
                                                  PassingRating+RushingAttempts+RushingYards+
                                                  RushingYardsPerAttempt+RushingTouchdowns+FumblesLost,
                                                  color=TeamIsHome))+
                     geom_point()+
                     geom_smooth(method="lm",se=F)+
                     theme(axis.text.x = element_text(angle=65, vjust=0.6))+facet_wrap(~TeamIsHome)

``` 


### 1.6.3 Relationship between FanatasyPoints with all predictors per team

and here......

```{r}
attach(QBCrossSectional)
QBCrossSectional %>% ggplot(aes(y=FantasyPoints,x=PassingCompletions+PassingAttempts+
                                                  PassingCompletionPercentage+
                                                  PassingYards+PassingYardsPerAttempt+
                                                  PassingTouchdowns+PassingInterceptions+
                                                  PassingRating+RushingAttempts+
                                                  RushingYards+RushingYardsPerAttempt+
                                                  RushingTouchdowns+FumblesLost,color=Team)) +
                     geom_point()+
                     geom_smooth(method="lm",se=F)+
                     theme(axis.text.x = element_text(angle=65, vjust=0.6))+facet_wrap(~Team)

``` 

### 1.6.4 Relationship between FantasyPoints and Individual predictors

    This makes sense - but we are actually trying to capture what the NEXT value for fantasy points is likely to be - these are all going to be highly correlated to the current week since the fantasy score is a linear combination of the predictors for any given week. We need to shift the data (and z-score it?) before running these charts.
    

```{r}
for (i in 7:19) {
  linear_plot <- QBCrossSectional %>% 
                 ggplot(aes_string(y="FantasyPoints",x=names(QBCrossSectional[i]))) + 
                 geom_point()+geom_smooth(method="lm",se=F)+
                 theme(axis.text.x = element_text(angle=65, vjust=0.6))+
                 theme_economist()
    print(linear_plot) 
  
    
  }

```


## Check for skewed predictors

```{r}
features_to_keep = c('PassingYards', 'PassingAttempts', 'PassingTouchdowns', 'PassingCompletions', 'PassingYardsPerAttempt')
for(f in features_to_keep){
  hist = QBCrossSectional %>% ggplot(aes_string(x=f))+
    geom_histogram()+
    theme(axis.text.x = element_text(angle=65, vjust=0.6))+
    theme_economist()
  print(hist)
}

#TODO: Check the normality by week?
```

Drop rushing yards per attempt

## Defensive stats

### Player Defensive Stats


```{r}

defensive_positions = getFootballData("https://fantasydata.com/FantasyStatsNFL/FantasyStats_Read?sort=FantasyPoints-desc&pageSize=4000&group=&filter=&filters.position=8&filters.team=&filters.teamkey=&filters.season=2017&filters.seasontype=1&filters.scope=2&filters.subscope=1&filters.redzonescope=&filters.scoringsystem=&filters.leaguetype=&filters.searchtext=&filters.week=&filters.startweek=1&filters.endweek=17&filters.minimumsnaps=&filters.teamaspect=&filters.stattype=&filters.exportType=&filters.desktop=&filter#s.dfsoperator=&filters.dfsslateid=&filters.dfsslategameid=&filters.dfsrosterslot=&filters.page=&filters.showfavs=&filters.posgroup=&filters.oddsstate=&filters.aggregatesc#ope=1&filters.rangescope=&filters.range=1")
defensive_positions$StatSummary = c(NULL)




```

##  Team Defensive Stats

```{r}

team_defense = getFootballData("https://fantasydata.com/FantasyStatsNFL/FantasyStats_Read?sort=FantasyPoints-desc&pageSize=4000&group=&filter=&filters.position=7&filters.team=&filters.teamkey=&filters.season=2017&filters.seasontype=1&filters.scope=2&filters.subscope=1&filters.redzonescope=&filters.scoringsystem=&filters.leaguetype=&filters.searchtext=&filters.week=&filters.startweek=1&filters.endweek=17&filters.minimumsnaps=&filters.teamaspect=&filters.stattype=&filters.exportType=&filters.desktop=&filters.dfsoperator=&filters.dfsslateid=&filters.dfsslategameid=&filters.dfsrosterslot=&filters.page=&filters.showfavs=&filters.posgroup=&filters.oddsstate=&filters.aggregatescope=1&filters.rangescope=&filters.range=1")
team_defense$StatSummary = c(NULL)
```

```{r}
defensive_players_by_week =  sqldf("SELECT 
                                      Week, 
                                      FantasyPosition,
                                      Count(*)
                                    FROM defensive_positions
                                    GROUP BY Week, FantasyPosition")

team_defense_by_week =  sqldf("SELECT 
                                      Week, 
                                      FantasyPosition,
                                      Count(*)
                                    FROM team_defense
                                    GROUP BY Week, FantasyPosition")

# this looks good some teams are on bye in certain weeks
```

## Do some averages for Defense before stitching on to QB data

```{r}
defensive_columns = c('Team', 'Week', 'TacklesForLoss', 'Sacks', 'QuarterbackHits', 'Interceptions', 'FumblesRecovered', 'Safeties', 'DefensiveTouchdowns', 'SoloTackles', 'AssistedTackles', 'SackYards', 'PassesDefended', 'FumblesForced', 'FantasyPoints', 'PointsAllowedByDefenseSpecialTeams')


team_defense = team_defense %>% select(defensive_columns) %>% rename('DefensiveFantasyPoints'='FantasyPoints') %>% group_by(Week) %>% mutate(WeeklyRank = dense_rank(DefensiveFantasyPoints)) 
attach(team_defense)
  
team_defense$Week = as.factor(team_defense$Week)  
  
sqldf("SELECT Week, Team, WeeklyRank FROM team_defense  WHERE Week IN(1, 2) ORDER BY Team, Week")
```

## Add defensive matchups

```{r}

QBCrossSectionalDefensiveOverlay = QBCrossSectional %>% left_join(team_defense, by = c('Week'='Week', 'Opponent'='Team' )) 

```

## Add some lag data for QB

### CumulativeVariables (these should definitley be combined into a weekly ranking)

```{r}

#Should we just train on second half of 2017?
QBCrossSectionalDefensiveOverlayCumulativePassYards = QBCrossSectionalDefensiveOverlay %>% group_by(PlayerID) %>% arrange(Week) %>%           mutate(CumulativeAveragePassingYards=cummean(PassingYards)
        , CumulativeAveragePassingTouchdowns=cummean(PassingTouchdowns)
        , CumulativeAverageCompletions = cummean(PassingCompletions) # not sure that completions matter much - most leagues don't reward them
        , CumulativeMaxPassingTouchdowns = cummax(PassingTouchdowns) 
        , CumulativeMaxPassingYards = cummax(PassingYards)
        , CumulativeMaxPassingAttempts = cummax(PassingAttempts)
        , CumulativeMaxPassingRating = cummax(PassingRating)
        , CumulativeMaxCompletions = cummax(PassingCompletions)
        , CumulativeMaxPassYardsPerAttempt = cummax(PassingYardsPerAttempt)
        , CumulativeMinPassingTouchdowns = cummin(PassingTouchdowns) 
        , CumulativeMinPassingYards = cummin(PassingYards)   #Let's get mins to capture downside risk
        , CumulativeMinPassingAttempts = cummin(PassingAttempts)
        , CumulativeMinPassingRating = cummin(PassingRating)
        , CumulativeMinCompletions = cummin(PassingCompletions)
        , CumulativeMinPassYardsPerAttempt = cummin(PassingYardsPerAttempt)
        , LastWeekQuarterBackRating = lag(PassingRating)
        , LastWeekQuarterPassingYards = lag(PassingYards)
        , LastWeekQuarterPassingTouchdowns = lag(PassingTouchdowns)
        , NextWeekFantasyPoints = lead(FantasyPoints) #Target Variable
        )

QBCrossSectionalDefensiveOverlayCumulativePassYards %>% filter(PlayerID == 6739) %>% write.csv('alex_smith.csv') #Alex Smith did indeed pass for 4042 yards on the season:)
```




## Do we want to convert any of these to factors?


