---
title: "Exploratory Data Analysis"
author: "Kevin Thompson, Sean Kennedy, Sachin Chavan"
date: "September 29, 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
# 1 : Load required libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(ggcorrplot)
library(visdat)
library(corrplot)
library(ggthemes)
library(sqldf)
library(rjson)
library(purrr)
library(gridExtra)
library(scales) # provides additional formatting options for ggplot objects
library(knitr)

```

# 2 : Load Quarterback Crosssectional data

```{r color="black"}
load("../analysis/data/.RData")
QBCrossSectional = getFootballData(URL) %>% filterUnwantedVariables() %>% filter(Position == 'QB')
summary(QBCrossSectional)
attach(QBCrossSectional)

```

Drop rushing yards per attempt

# 3 : Defensive stats 

    ##  3.1 : Team Defensive Stats

```{r}

team_defense = getFootballData("https://fantasydata.com/FantasyStatsNFL/FantasyStats_Read?sort=FantasyPoints-desc&pageSize=4000&group=&filter=&filters.position=7&filters.team=&filters.teamkey=&filters.season=2017&filters.seasontype=1&filters.scope=2&filters.subscope=1&filters.redzonescope=&filters.scoringsystem=&filters.leaguetype=&filters.searchtext=&filters.week=&filters.startweek=1&filters.endweek=17&filters.minimumsnaps=&filters.teamaspect=&filters.stattype=&filters.exportType=&filters.desktop=&filters.dfsoperator=&filters.dfsslateid=&filters.dfsslategameid=&filters.dfsrosterslot=&filters.page=&filters.showfavs=&filters.posgroup=&filters.oddsstate=&filters.aggregatescope=1&filters.rangescope=&filters.range=1")
team_defense$StatSummary = c(NULL)

defensive_columns = c('Team', 'Week', 'TacklesForLoss', 'Sacks', 'QuarterbackHits', 'Interceptions', 'FumblesRecovered', 'Safeties', 'DefensiveTouchdowns', 'SoloTackles', 'AssistedTackles', 'SackYards', 'PassesDefended', 'FumblesForced', 'FantasyPoints', 'PointsAllowedByDefenseSpecialTeams')


team_defense = team_defense %>% dplyr::select(defensive_columns) %>% rename('DefensiveFantasyPoints'='FantasyPoints') 

attach(team_defense)
```

    ## 3.5 : Add defensive matchups

```{r}


#This only adds the current weeks matchup - we need next week's matchup as a target

team_defense_custom = sqldf("SELECT Team
                                    ,Week 
                                    ,(Sacks * 4
                                    +QuarterbackHits * 3
                                    +Interceptions * 7
                                    +SackYards * 2) as PassingDefense
                                    ,PointsAllowedByDefenseSpecialTeams
                                    FROM team_defense")
#scheme: weight interceptions, qbsacks, quarterbackhits, passesdefended

plot(team_defense_custom$PassingDefense, team_defense_custom$PointsAllowedByDefenseSpecialTeams)

team_defensive_rankings = team_defense_custom %>% 
                              #filter(Week > 1 & Week < 17)
                              group_by(Team) %>%
                              arrange(Week) %>%
                              mutate(
                                AvgPassDefense = cumsum(PassingDefense)/Week #Account for bye week....
                              )
QBCrossSectional$Week = as.numeric(QBCrossSectional$Week)

hist(team_defensive_rankings$AvgPassDefense)

sqldf("SELECT * FROM team_defensive_rankings WHERE Team = 'LAC'")

```

    ## 3.6 : Add some lag data for QB

    ### CumulativeVariables (these should definitley be combined into a weekly ranking)

```{r}

QBCrossSectionalCumulativePassYards = QBCrossSectional %>% group_by(PlayerID)%>% filter(n() >= 8) %>% arrange(Week) %>%  
   mutate(CumulativeAveragePassingYards=cummean(PassingYards)
        , CumulativeAveragePassingTouchdowns=cummean(PassingTouchdowns)
        , CumulativeAveragePassingInterceptions=cummean(PassingInterceptions)
        , CumulativeAveragePassingRating=cummean(PassingRating)
        , CumulativeAverageCompletions = cummean(PassingCompletions) # not sure that completions matter much - most leagues don't reward them
        , CumulativeAverageCompletionPercentage = cummean(PassingCompletionPercentage)
        , CumulativeMaxPassingTouchdowns = cummax(PassingTouchdowns) 
        , CumulativeMaxPassingYards = cummax(PassingYards)
        , CumulativeMaxPassingAttempts = cummax(PassingAttempts)
        , CumulativeMaxPassingRating = cummax(PassingRating)
        , CumulativeMaxCompletions = cummax(PassingCompletions)
        , CumulativeMaxPassYardsPerAttempt = cummax(PassingYardsPerAttempt)
        , CumulativeMinPassingTouchdowns = cummin(PassingTouchdowns) 
        , CumulativeMinPassingYards = cummin(PassingYards)   #Let's get mins to capture downside risk
        , CumulativeMinPassingAttempts = cummin(PassingAttempts)
        , CumulativeMinPassingRating = cummin(PassingRating)
        , CumulativeMinCompletions = cummin(PassingCompletions)
        , CumulativeMinPassYardsPerAttempt = cummin(PassingYardsPerAttempt)
        , NextWeekFantasyPoints = lead(FantasyPoints) #Target Variable
        , NextOpponent = lead(Opponent)
        #, NextWeekDefensiveMatchup = lag(WeeklyRank) #Has to be last week's team ranking - not this week
      )


QBCrossSectionalDefensiveOverlayCumulativePassYards = QBCrossSectionalCumulativePassYards %>% left_join(team_defensive_rankings, by = c('Week'='Week', 'NextOpponent'='Team')) 

#QBCrossSectionalDefensiveOverlayCumulativePassYards %>% filter(PlayerID == 6739) %>% write.csv('alex_smith.csv') #Alex Smith did indeed pass for 4042 yards on the season:)
attach(QBCrossSectionalDefensiveOverlayCumulativePassYards)

summary(QBCrossSectionalDefensiveOverlayCumulativePassYards)
```

     ## 3.7 : Create indepenent structure for EDA from QBCrossSectional Data

```{r}

eda_base <- QBCrossSectional %>% group_by(Week) %>% 
                        select(Week,
                               FantasyPoints,
                               PassingCompletions,
                               PassingAttempts,
                               PassingCompletionPercentage,
                               PassingYards, 
                               PassingYardsPerAttempt,
                               PassingTouchdowns,
                               PassingInterceptions,
                               PassingRating,
                               RushingAttempts,
                               RushingYards,
                               RushingYardsPerAttempt,
                               RushingTouchdowns,
                               FumblesLost
                               )

eda_base$Week <- as.factor(eda_base$Week)

```

     ## 3.8 : Create indepenent structure for derived features for EDA
              from QBCrossSectionalDefensiveOverlayCumulativePassYards Data

```{r}

eda_derived <- QBCrossSectionalDefensiveOverlayCumulativePassYards %>% group_by(Week)%>% 
               select(Week,
                      FantasyPoints,                        
                      CumulativeAveragePassingYards,        
                      CumulativeAveragePassingTouchdowns,   
                      CumulativeAveragePassingInterceptions,
                      CumulativeAveragePassingRating,       
                      CumulativeAverageCompletions,         
                      CumulativeAverageCompletionPercentage,
                      CumulativeMaxPassingTouchdowns,       
                      CumulativeMaxPassingYards,            
                      CumulativeMaxPassingAttempts,        
                      CumulativeMaxPassingRating,           
                      CumulativeMaxCompletions,             
                      CumulativeMaxPassYardsPerAttempt,     
                      CumulativeMinPassingTouchdowns,       
                      CumulativeMinPassingYards,           
                      CumulativeMinPassingAttempts,        
                      CumulativeMinPassingRating,          
                      CumulativeMinCompletions,            
                      CumulativeMinPassYardsPerAttempt) 
eda_derived$Week <- as.factor(eda_derived$Week)
```

# 4 : Structures QBCrossSectional (Quarterbacks)

```{r}
summary(QBCrossSectional)
```

# 5 : struture with Derived features

```{r}
summary(eda_derived)
```



# 6 : QBCrossSectional - Check Missing Data 

```{r}
vis_dat(QBCrossSectional)
```


# 7 : Corrrelogram 

    ## 7.1 : Base Feature Set
        
```{r}
corr <- round(cor(eda_base[-1]), 1)

ggcorrplot(corr, hc.order = TRUE, 
           type = "full", 
           lab = TRUE, 
           lab_size = 1.5, 
           method="square", 
           colors = c("tomato2", "white", "springgreen3"), 
           title  ="Correlogram of Base QB features", 
           tl.cex = 7,pch=2,pch.col =3,show.diag = T,
           ggtheme=theme_classic)
```
      
    ## 7.2 : Derived Feature Set

```{r}
corr <- round(cor(eda_derived[-1]), 1)

ggcorrplot(corr, hc.order = TRUE, 
           type = "full", 
           lab = TRUE, 
           lab_size = 1.5, 
           method="square", 
           colors = c("tomato2", "white", "springgreen3"), 
           title  = "Correlogram of Derived QB features", 
           tl.cex = 7,pch=2,pch.col =3,show.diag = T,
           ggtheme=theme_classic)
```

# 8 : Distributions

    ## 8.1 Density plot for Fantasypoints is approxmiately Normal

```{r}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
target <- QBCrossSectional$FantasyPoints
plot(density(target), main="CrossSectional Dataset: FantasyPoints", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(target), 2)))
polygon(density(target), col="red")

```

    # 8.2 : Boxplots - Target and Individual Predictor Bevavior for per Team

    ## 8.2.1 Base Feature set

```{r}

eda_base %>% ggplot(aes(y=FantasyPoints,x=Week,fill=Week,group=Week))+
                     geom_boxplot(show.legend = FALSE)+
                     xlab("Week")+ylab("FantasyPoints")+
                     labs(title="Fantasyfootball", 
                     subtitle="Weekly fantasypoints", 
                     aption="Source: Fantasyfootball")+
                     theme_wsj()+
                     theme(plot.title    = element_text(size = rel(0.5)),
                           plot.subtitle = element_text(size = rel(0.5)),
                           axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                           axis.title    = element_text(size = rel(0.5)),
                           legend.position  = "right",
                           legend.direction ="vertical",
                           legend.title = element_text(size = rel(0.5)))   

for (i in 2:15) {

ggplotp <- eda_base %>% 
           ggplot(aes_string(y=names(eda_base[i]),x="Week",fill="Week",group="Week"))+
           geom_boxplot(show.legend = FALSE)+
           xlab("Week")+ylab(names(eda_base[i]))+
           labs(title="Fantasyfootball", 
           subtitle=names(eda_base[i]), 
           aption="Source: Fantasyfootball")+
           theme_wsj()+
           theme(plot.title    = element_text(size = rel(0.5)),
                 plot.subtitle = element_text(size = rel(0.5)),
                 axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                 axis.title    = element_text(size = rel(0.5)),
                 legend.position  = "right",
                 legend.direction ="vertical",
                 legend.title = element_text(size = rel(0.5)))   
  

  print(ggplotp)
}



```

    ## 8.2.3 Boxplots Derived Feature set

```{r}

eda_derived %>% ggplot(aes(y=FantasyPoints,x=Week,fill=Week,group=Week))+
                      geom_boxplot(show.legend = FALSE)+
                      xlab("Week")+ylab("FantasyPoints")+
                      labs(title="Fantasyfootball", 
                      subtitle="Weekly fantasypoints", 
                      aption="Source: Fantasyfootball")+
                      theme_wsj()+
                      theme(plot.title    = element_text(size = rel(0.5)),
                            plot.subtitle = element_text(size = rel(0.5)),
                            axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                            axis.title    = element_text(size = rel(0.5)),
                            legend.position  = "right",
                            legend.direction ="vertical",
                            legend.title = element_text(size = rel(0.5)))   

for (i in 2:20) {

ggplotp <- eda_derived %>% 
           ggplot(aes_string(y=names(eda_derived[i]),x="Week",fill="Week",group="Week"))+
           geom_boxplot(show.legend = FALSE)+
           xlab("Week")+ylab(names(eda_derived[i]))+
           labs(title="Fantasyfootball", 
           subtitle=names(eda_derived[i]), 
           aption="Source: Fantasyfootball")+
           theme_wsj()+
           theme(plot.title    = element_text(size = rel(0.5)),
                 plot.subtitle = element_text(size = rel(0.5)),
                 axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                 axis.title    = element_text(size = rel(0.5)),
                 legend.position  = "right",
                 legend.direction ="vertical",
                 legend.title = element_text(size = rel(0.5)))   
  

  print(ggplotp)
}


```

    ## 8.3 Violin plots Derived Feature set

```{r}

eda_derived %>% ggplot(aes(y=FantasyPoints,x=Week,fill=Week,group=Week))+
                      geom_violin(show.legend = FALSE)+
                      xlab("Week")+ylab("FantasyPoints")+
                      labs(title="Fantasyfootball", 
                      subtitle="Weekly fantasypoints", 
                      aption="Source: Fantasyfootball")+
                      theme_wsj()+
                      theme(plot.title    = element_text(size = rel(0.5)),
                            plot.subtitle = element_text(size = rel(0.5)),
                            axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                            axis.title    = element_text(size = rel(0.5)),
                            legend.position  = "right",
                            legend.direction ="vertical",
                            legend.title = element_text(size = rel(0.5)))   

for (i in 2:20) {

ggplotp <- eda_derived %>% 
           ggplot(aes_string(y=names(eda_derived[i]),x="Week",fill="Week",group="Week"))+
           geom_violin(show.legend = FALSE)+
           xlab("Week")+ylab(names(eda_derived[i]))+
           labs(title="Fantasyfootball", 
           subtitle=names(eda_derived[i]), 
           aption="Source: Fantasyfootball")+
           theme_wsj()+
           theme(plot.title    = element_text(size = rel(0.5)),
                 plot.subtitle = element_text(size = rel(0.5)),
                 axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                 axis.title    = element_text(size = rel(0.5)),
                 legend.position  = "right",
                 legend.direction ="vertical",
                 legend.title = element_text(size = rel(0.5)))   
  

  print(ggplotp)
}
```

    ### Check for skewed predictors

    ## 8.4 : Histograms - Normality by Week
    
    ### 8.4.1 : Base feature Set

```{r}
features_to_keep = c('PassingYards', 'PassingAttempts', 'PassingTouchdowns', 'PassingCompletions', 'PassingYardsPerAttempt')

for(f in features_to_keep){
  hist = eda_base %>% ggplot(aes_string(x=f,fill="Week"))+
                      geom_histogram(bins=30,show.legend = FALSE)+
                      theme_wsj()+
                      theme(plot.title    = element_text(size = rel(0.5)),
                            plot.subtitle = element_text(size = rel(0.5)),
                            axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                            axis.title    = element_text(size = rel(0.5)),
                            legend.position  = "right",
                            legend.direction ="vertical",
                            legend.title = element_text(size = rel(0.5)))   
  print(hist)
}

```

    ### 8.4.2 : Dervived feature Set

```{r}
features_to_keep = names(eda_derived[-1])

for(f in features_to_keep){
  hist = eda_derived %>% ggplot(aes_string(x=f,fill="Week"))+
                      geom_histogram(bins=30,show.legend = FALSE)+
                      theme_wsj()+    
                      theme(plot.title    = element_text(size = rel(0.5)),
                            plot.subtitle = element_text(size = rel(0.5)),
                            axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                            axis.title    = element_text(size = rel(0.5)),
                            legend.position  = "right",
                            legend.direction ="vertical",
                            legend.title = element_text(size = rel(0.5)))   
  print(hist)
}

```

    ## 8.5 Bar plots Derived Feature set


```{r}

for(p in 2:20){
    bar_plot  <- eda_derived %>% 
                 ggplot(aes_string(x="Week",
                                   y=names(eda_derived[p]),
                                   fill="Week",group="Week"))+
                 ggtitle(names(eda_derived[p]))+
                 geom_bar(stat="identity", width=.5, show.legend = FALSE)+
                 xlab("Week")+
                 ylab(names(eda_derived[p]))+  
                 labs(title="Fantasyfootball", 
                      subtitle=names(eda_derived[p]), 
                      aption="Source: Fantasyfootball")+
                  theme_wsj()+
                  theme(plot.title    = element_text(size = rel(0.5)),
                        plot.subtitle = element_text(size = rel(0.5)),
                        axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                        axis.title    = element_text(size = rel(0.5)),
                        legend.position  = "right",
                        legend.direction ="vertical",
                        legend.title = element_text(size = rel(0.5)))   
           print(bar_plot)
}

```


# 9 :  Scatterplots

    We are trying to capture what the NEXT value for fantasy points is likely to be - these are all going to be highly correlated to the
    current week since the fantasy score is a linear combination of the predictors for any given week. 
    We need to shift the data before running these charts.
    
    ## 9.1 : Scatterplots for base features

```{r}
for (i in 2:15) {
  base_scatter <- eda_base %>% 
                  ggplot(aes_string(y="FantasyPoints",x=names(eda_base[i]),color="as.numeric(Week)")) + 
                  geom_point()+geom_smooth(method="lm",se=F)+
                  labs(title="Fantasyfootball", 
                       subtitle=names(eda_base[i]), 
                       aption="Source: Fantasyfootball")+
                  labs(color="Week")+
                  theme_wsj()+
                  theme(plot.title    = element_text(size = rel(0.5)),
                        plot.subtitle = element_text(size = rel(0.5)),
                        axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                        axis.title    = element_text(size = rel(0.5)),
                        legend.position  = "right",
                        legend.direction ="vertical",
                        legend.title = element_text(size = rel(0.5))
                       )
    print(base_scatter) 
  }
```

    ## 9.2 : Scatterplots for Derived features

```{r}

for (i in 2:20) {
  derived_scatter <- eda_derived %>% 
                     ggplot(aes_string(y="FantasyPoints",x=names(eda_derived[i]),color="as.numeric(Week)"))+ 
                     geom_point()+
                     geom_smooth(method="lm",se=F)+
                     xlab(names(eda_derived[i]))+
                     ylab("FantasyPoints")+  
                     labs(title="Fantasyfootball", 
                          subtitle=names(eda_derived[i]), 
                          aption="Source: Fantasyfootball")+
                     labs(color="Week")+
                     theme_wsj()+
                     theme(plot.title    = element_text(size = rel(0.5)),
                           plot.subtitle = element_text(size = rel(0.5)),
                           axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                           axis.title    = element_text(size = rel(0.5)),
                           legend.position  = "right",
                           legend.direction ="vertical",
                           legend.title = element_text(size = rel(0.5))
                          )
    print(derived_scatter) 
}

```

# 10 : Line plots
  ## 10.1 Base Features

```{r}

line_ds <- eda_base %>% group_by(Week) %>% arrange(Week)

for(p in 2:15){

   line_plot  <-  line_ds %>% 
                  ggplot(aes_string(x="as.numeric(Week)",y=names(line_ds[p])))+
                  ggtitle(names(line_ds[p]))+
                  geom_line(show.legend = FALSE)+
                  xlab("Week")+
                  ylab(names(line_ds[p]))+  
                  labs(title="Fantasyfootball", 
                       subtitle=names(line_ds[p]), 
                       aption="Source: Fantasyfootball")+
                  theme_wsj()+
                  theme(plot.title    = element_text(size = rel(0.5)),
                        plot.subtitle = element_text(size = rel(0.5)),
                        axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                        axis.title    = element_text(size = rel(0.5)),
                        legend.position  = "right",
                        legend.direction ="vertical",
                        legend.title = element_text(size = rel(0.5)))   
       
   print(line_plot)
}

```
  
  
  ## 10.2 Derived Features
  
```{r}

line_ds <- eda_derived %>% group_by(Week) %>% arrange(Week)

for(p in 2:20){

   line_plot  <-  line_ds %>% 
                  ggplot(aes_string(x="as.numeric(Week)",y=names(line_ds[p])))+
                  ggtitle(names(line_ds[p]))+
                  geom_line(show.legend = FALSE)+
                  xlab("Week")+
                  ylab(names(line_ds[p]))+  
                  labs(title="Fantasyfootball", 
                       subtitle=names(line_ds[p]), 
                       aption="Source: Fantasyfootball")+
                  theme_wsj()+
                  theme(plot.title    = element_text(size = rel(0.5)),
                        plot.subtitle = element_text(size = rel(0.5)),
                        axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                        axis.title    = element_text(size = rel(0.5)),
                        legend.position  = "right",
                        legend.direction ="vertical",
                        legend.title = element_text(size = rel(0.5)))   
       
   print(line_plot)
}

```




