---
title: "Exploratory Data Analysis"
author: "Kevin Thompson, Sean Kennedy, Sachin Chavan"
date: "September 29, 2019"
output:
  html_document: default
  pdf_document: default
---
# 1 : Load required libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(ggcorrplot)
library(visdat)
library(corrplot)
library(ggthemes)
library(sqldf)
library(rjson)
library(purrr)
library(gridExtra)
library(scales) # provides additional formatting options for ggplot objects
library(knitr)

```

# 2 : Load Quarterback Crosssectional data

```{r color="black"}
load("../analysis/data/.RData")
QBCrossSectional$Week <- as.factor(QBCrossSectional$Week)
attach(QBCrossSectional)
```

Drop rushing yards per attempt

# 3 : Defensive stats 

    ## 3.1 : Player Defensive Stats


```{r}

defensive_positions = getFootballData("https://fantasydata.com/FantasyStatsNFL/FantasyStats_Read?sort=FantasyPoints-desc&pageSize=4000&group=&filter=&filters.position=8&filters.team=&filters.teamkey=&filters.season=2017&filters.seasontype=1&filters.scope=2&filters.subscope=1&filters.redzonescope=&filters.scoringsystem=&filters.leaguetype=&filters.searchtext=&filters.week=&filters.startweek=1&filters.endweek=17&filters.minimumsnaps=&filters.teamaspect=&filters.stattype=&filters.exportType=&filters.desktop=&filter#s.dfsoperator=&filters.dfsslateid=&filters.dfsslategameid=&filters.dfsrosterslot=&filters.page=&filters.showfavs=&filters.posgroup=&filters.oddsstate=&filters.aggregatesc#ope=1&filters.rangescope=&filters.range=1")

defensive_positions$StatSummary = c(NULL)




```

    ##  3.2 : Team Defensive Stats

```{r}

team_defense = getFootballData("https://fantasydata.com/FantasyStatsNFL/FantasyStats_Read?sort=FantasyPoints-desc&pageSize=4000&group=&filter=&filters.position=7&filters.team=&filters.teamkey=&filters.season=2017&filters.seasontype=1&filters.scope=2&filters.subscope=1&filters.redzonescope=&filters.scoringsystem=&filters.leaguetype=&filters.searchtext=&filters.week=&filters.startweek=1&filters.endweek=17&filters.minimumsnaps=&filters.teamaspect=&filters.stattype=&filters.exportType=&filters.desktop=&filters.dfsoperator=&filters.dfsslateid=&filters.dfsslategameid=&filters.dfsrosterslot=&filters.page=&filters.showfavs=&filters.posgroup=&filters.oddsstate=&filters.aggregatescope=1&filters.rangescope=&filters.range=1")
team_defense$StatSummary = c(NULL)


```

    ## 3.3 : Defensive Players and team defense by week

```{r}
defensive_players_by_week =  sqldf("SELECT 
                                      Week, 
                                      FantasyPosition,
                                      Count(*)
                                    FROM defensive_positions
                                    GROUP BY Week, FantasyPosition")

team_defense_by_week =  sqldf("SELECT 
                                      Week, 
                                      FantasyPosition,
                                      Count(*)
                                    FROM team_defense
                                    GROUP BY Week, FantasyPosition")

# this looks good some teams are on bye in certain weeks
```

    ## 3.4 : Do some averages for Defense before stitching on to QB data

```{r}
defensive_columns = c('Team', 'Week', 'TacklesForLoss', 'Sacks', 'QuarterbackHits', 'Interceptions', 'FumblesRecovered', 'Safeties', 'DefensiveTouchdowns', 'SoloTackles', 'AssistedTackles', 'SackYards', 'PassesDefended', 'FumblesForced', 'FantasyPoints', 'PointsAllowedByDefenseSpecialTeams')


team_defense = team_defense %>% select(defensive_columns) %>% rename('DefensiveFantasyPoints'='FantasyPoints') %>% group_by(Week) %>%
               mutate(WeeklyRank = dense_rank(DefensiveFantasyPoints))
#               mutate(PlayerName = paste(FirstName," ",LastName))
attach(team_defense)
  
team_defense$Week = as.factor(team_defense$Week)  
  
sqldf("SELECT Week, Team, WeeklyRank FROM team_defense  WHERE Week IN(1, 2) ORDER BY Team, Week")
```

    ## 3.5 : Add defensive matchups

```{r}

QBCrossSectionalDefensiveOverlay = QBCrossSectional %>% left_join(team_defense, by = c('Week'='Week', 'Opponent'='Team')) 

```

    ## 3.6 : Add some lag data for QB

    ### CumulativeVariables (these should definitley be combined into a weekly ranking)

```{r}

QBCrossSectionalDefensiveOverlayCumulativePassYards = QBCrossSectionalDefensiveOverlay %>% group_by(PlayerID) %>% arrange(Week) %>%           mutate(CumulativeAveragePassingYards=cummean(PassingYards)
        , CumulativeAveragePassingTouchdowns=cummean(PassingTouchdowns)
        , CumulativeAveragePassingRating=cummean(PassingRating)
        , CumulativeAverageCompletions = cummean(PassingCompletions) # not sure that completions matter much - most leagues don't reward them
        , CumulativeAverageCompletionPercentage = cummean(PassingCompletionPercentage)
        , CumulativeMaxPassingTouchdowns = cummax(PassingTouchdowns) 
        , CumulativeMaxPassingYards = cummax(PassingYards)
        , CumulativeMaxPassingAttempts = cummax(PassingAttempts)
        , CumulativeMaxPassingRating = cummax(PassingRating)
        , CumulativeMaxCompletions = cummax(PassingCompletions)
        , CumulativeMaxPassYardsPerAttempt = cummax(PassingYardsPerAttempt)
        , CumulativeMinPassingTouchdowns = cummin(PassingTouchdowns) 
        , CumulativeMinPassingYards = cummin(PassingYards)   #Let's get mins to capture downside risk
        , CumulativeMinPassingAttempts = cummin(PassingAttempts)
        , CumulativeMinPassingRating = cummin(PassingRating)
        , CumulativeMinCompletions = cummin(PassingCompletions)
        , CumulativeMinPassYardsPerAttempt = cummin(PassingYardsPerAttempt)
        #, LastWeekQuarterBackRating = lag(PassingRating) by leading target variable to t+1, the current values are implicitly lagged
        #, LastWeekPassingYards = lag(PassingYards)
        #, LastWeekPassingTouchdowns = lag(PassingTouchdowns)
        , NextWeekFantasyPoints = lead(FantasyPoints) #Target Variable
        , NextWeekDefensiveMatchup = lag(WeeklyRank) #Has to be last week's team ranking - not this week
      )

QBCrossSectionalDefensiveOverlayCumulativePassYards %>% filter(PlayerID == 6739) %>% write.csv('alex_smith.csv') #Alex Smith did indeed pass for 4042 yards on the season:)
final_ds <- QBCrossSectionalDefensiveOverlayCumulativePassYards
```



# 4 : Structures QBCrossSectional (Quarterbacks)

```{r}
summary(QBCrossSectional)
```

# 5 : struture with Derived features

```{r}
summary(final_ds)
```



# 6 : QBCrossSectional - Check Missing Data 

```{r}
vis_dat(QBCrossSectional)
```


# 7 : Corrrelogram 

    ## 7.1 : Base Feature Set
        
```{r}
corr <- round(cor(QBCrossSectional[-c(1:6,21)]), 1)

ggcorrplot(corr, hc.order = TRUE, 
           type = "full", 
           lab = TRUE, 
           lab_size = 1.5, 
           method="square", 
           colors = c("tomato2", "white", "springgreen3"), 
           title  ="Correlogram of Quarterbacks", 
           tl.cex = 7,pch=2,pch.col =3,show.diag = T,
           ggtheme=theme_classic)
```
      
    ## 7.2 : Derived Feature Set

```{r}
final_cor_ds <- final_ds %>% select(PlayerID,FantasyPoints,names(final_ds[37:51]))

corr <- round(cor(final_cor_ds[-1]), 1)

ggcorrplot(corr, hc.order = TRUE, 
           type = "full", 
           lab = TRUE, 
           lab_size = 1.5, 
           method="square", 
           colors = c("tomato2", "white", "springgreen3"), 
           title  = "Correlogram of Cumulative Averages", 
           tl.cex = 7,pch=2,pch.col =3,show.diag = T,
           ggtheme=theme_classic)
```

# 8 : Distributions

    ## 8.1 Density plot for Fantasypoints is approxmiately Normal

```{r}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
target <- QBCrossSectional$FantasyPoints
plot(density(target), main="CrossSectional Dataset: FantasyPoints", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(target), 2)))
polygon(density(target), col="red")

```

    # 8.2 : Boxplots - Target and Individual Predictor Bevavior for per Team

    ## 8.2.1 Base Feature set

```{r}

QBCrossSectional %>% ggplot(aes(y=FantasyPoints,x=Week,fill=Week,group=Week))+
                     geom_boxplot(show.legend = FALSE)+
                     xlab("Week")+ylab("FantasyPoints")+
                     labs(title="Fantasyfootball", 
                     subtitle="Weekly fantasypoints", 
                     aption="Source: Fantasyfootball")+
                     theme_wsj()+
                     theme(plot.title    = element_text(size = rel(0.5)),
                           plot.subtitle = element_text(size = rel(0.5)),
                           axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                           axis.title    = element_text(size = rel(0.5)),
                           legend.position  = "right",
                           legend.direction ="vertical",
                           legend.title = element_text(size = rel(0.5)))   

for (i in 7:19) {

ggplotp <- QBCrossSectional %>% 
           ggplot(aes_string(y=names(QBCrossSectional[i]),x="Week",fill="Week",group="Week"))+
           geom_boxplot(show.legend = FALSE)+
           xlab("Week")+ylab(names(QBCrossSectional[i]))+
           labs(title="Fantasyfootball", 
           subtitle=names(QBCrossSectional[i]), 
           aption="Source: Fantasyfootball")+
           theme_wsj()+
           theme(plot.title    = element_text(size = rel(0.5)),
                 plot.subtitle = element_text(size = rel(0.5)),
                 axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                 axis.title    = element_text(size = rel(0.5)),
                 legend.position  = "right",
                 legend.direction ="vertical",
                 legend.title = element_text(size = rel(0.5)))   
  

  print(ggplotp)
}



```

    ## 8.2.3 Derived Feature set

```{r}

for(p in 37:51){
  for (player in seq(1,length(unique(final_ds$PlayerID)),by=20)){
       a_size <- player + 19
       players_ds <- final_ds %>% filter(PlayerID %in% unique(final_ds$PlayerID)[player:a_size])
       
       b_plot  <-  players_ds %>% 
                   ggplot(aes_string(x="Week",y=names(players_ds[p]),fill="Week",group="Week"))+
                   ggtitle(names(players_ds[p]))+
                   geom_boxplot(show.legend = FALSE)+
                   xlab("Week")+
                   ylab(names(players_ds[p]))+  
                   labs(title="Fantasyfootball", 
                        subtitle=names(players_ds[p]), 
                        aption="Source: Fantasyfootball")+
                  theme_economist(base_size=8)+
                  theme(axis.text.y = element_text(size=5))+
                  theme(axis.text.x = element_text(angle=65, vjust=0.6))
      print(b_plot)
   }
}


```

    ## 8.3 Violin plots Derived Feature set

```{r}
for(p in 37:51){
  #print(paste("violin plots for ",names(players_ds[p])))
   for (player in seq(1,length(unique(final_ds$PlayerID)),by=20)){
       a_size <- player + 19
       players_ds <- final_ds %>% filter(PlayerID %in% unique(final_ds$PlayerID)[player:a_size])
       
       v_plot  <-  players_ds %>% 
                   ggplot(aes_string(x="Week",y=names(players_ds[p]),fill="Week",group="Week"))+
                   ggtitle(names(players_ds[p]))+
                   geom_violin(show.legend = FALSE)+
                   xlab("Week")+
                   ylab(names(players_ds[p]))+  
                   labs(title="Fantasyfootball", 
                        subtitle=names(players_ds[p]), 
                        aption="Source: Fantasyfootball")+
                  theme_economist(base_size=8)+
                  theme(axis.text.y = element_text(size=5))+
                  theme(axis.text.x = element_text(angle=65, vjust=0.6))
      print(v_plot)
   }
}

```

    ### Check for skewed predictors

    ## 8.4 : Histograms - Normality by Week
    
    ### 8.4.1 : Base feature Set

```{r}
features_to_keep = c('PassingYards', 'PassingAttempts', 'PassingTouchdowns', 'PassingCompletions', 'PassingYardsPerAttempt')

for(f in features_to_keep){
  hist = QBCrossSectional %>% ggplot(aes_string(x=f,fill="Week"))+
                              geom_histogram(bins=30,show.legend = FALSE)+
                              theme_economist()+
                              theme(axis.text.x = element_text(angle = 45, hjust = 1))
  print(hist)
}

```

    ### 8.4.2 : Dervived feature Set

```{r}
features_to_keep = names(final_ds[37:51])

for(f in features_to_keep){
  hist = final_ds %>% ggplot(aes_string(x=f,fill="Week"))+
                      geom_histogram(bins=30,show.legend = FALSE)+
                      theme_classic()+
                      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  print(hist)
}

```

    ## 8.5 Bar plots Derived Feature set


```{r}

for(p in 37:51){
    bar_plot  <- final_ds %>% 
                 ggplot(aes_string(x="Week",
                                   y=names(final_ds[p]),
                                   fill="Week",group="Week"))+
                 ggtitle(names(final_ds[p]))+
                 geom_bar(stat="identity", width=.5, show.legend = FALSE)+
                 xlab("Week")+
                 ylab(names(final_ds[p]))+  
                 labs(title="Fantasyfootball", 
                      subtitle=names(final_ds[p]), 
                      aption="Source: Fantasyfootball")+
                  theme_wsj()+
                  theme(plot.title    = element_text(size = rel(0.5)),
                        plot.subtitle = element_text(size = rel(0.5)),
                        axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                        axis.title    = element_text(size = rel(0.5)),
                        legend.position  = "right",
                        legend.direction ="vertical",
                        legend.title = element_text(size = rel(0.5)))   
           print(bar_plot)
}

bar_ds <- final_ds %>%  group_by(Week)

```


# 9 :  Scatterplots

    We are trying to capture what the NEXT value for fantasy points is likely to be - these are all going to be highly correlated to the
    current week since the fantasy score is a linear combination of the predictors for any given week. 
    We need to shift the data before running these charts.
    
    ## 9.1 : Scatterplots for base features

```{r}
for (i in 7:19) {
  base_scatter <- QBCrossSectional %>% 
                  ggplot(aes_string(y="FantasyPoints",x=names(QBCrossSectional[i]),color="as.numeric(Week)")) + 
                  geom_point()+geom_smooth(method="lm",se=F)+
                  labs(title="Fantasyfootball", 
                       subtitle=names(final_ds[i]), 
                       aption="Source: Fantasyfootball")+
                  labs(color="Week")+
                  theme_wsj()+
                  theme(plot.title    = element_text(size = rel(0.5)),
                        plot.subtitle = element_text(size = rel(0.5)),
                        axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                        axis.title    = element_text(size = rel(0.5)),
                        legend.position  = "right",
                        legend.direction ="vertical",
                        legend.title = element_text(size = rel(0.5))
                       )
    print(base_scatter) 
  }
```

    ## 9.2 : Scatterplots for Derived features

```{r}

for (i in 37:51) {
  derived_scatter <- final_ds %>% 
                     ggplot(aes_string(y="FantasyPoints",x=names(final_ds[i]),color="as.numeric(Week)"))+ 
                     geom_point()+
                     geom_smooth(method="lm",se=F)+
                     xlab(names(final_ds[i]))+
                     ylab("FantasyPoints")+  
                     labs(title="Fantasyfootball", 
                          subtitle=names(final_ds[i]), 
                          aption="Source: Fantasyfootball")+
                     labs(color="Week")+
                     theme_wsj()+
                     theme(plot.title    = element_text(size = rel(0.5)),
                           plot.subtitle = element_text(size = rel(0.5)),
                           axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                           axis.title    = element_text(size = rel(0.5)),
                           legend.position  = "right",
                           legend.direction ="vertical",
                           legend.title = element_text(size = rel(0.5))
                          )
    print(derived_scatter) 
}

```

# 10 : Line plots

```{r}

line_ds <- final_ds %>% group_by(Week) %>% arrange(Week)

for(p in 37:51){

   line_plot  <-  line_ds %>% 
                  ggplot(aes_string(x="as.numeric(Week)",y=names(line_ds[p])))+
                  ggtitle(names(line_ds[p]))+
                  geom_line(show.legend = FALSE)+
                  xlab("Week")+
                  ylab(names(line_ds[p]))+  
                  labs(title="Fantasyfootball", 
                       subtitle=names(line_ds[p]), 
                       aption="Source: Fantasyfootball")+
                  theme_wsj()+
                  theme(plot.title    = element_text(size = rel(0.5)),
                        plot.subtitle = element_text(size = rel(0.5)),
                        axis.text.x   = element_text(angle=65, vjust=0.6,size=1),
                        axis.title    = element_text(size = rel(0.5)),
                        legend.position  = "right",
                        legend.direction ="vertical",
                        legend.title = element_text(size = rel(0.5)))   
       
   print(line_plot)
}

```




